---
title: "uncertainty"
author: "Harrison Leduc"
date: "2024-04-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Data Import and Insight
```{r}
library(dplyr)
library(sf)
library(terra)
library(parallel)

data_dir <- here::here("external/afcdata")

flds <- sfarrow::st_read_parquet(
  file.path(data_dir, "mapped_fields_final.parquet"))

ctlg <- readr::read_csv(
  file.path(data_dir, "label_catalog_allclasses.csv"))
```

# Spatial Features/Characteristics
```{r}
library(ggplot2)

polybox <- function(x, y, w = 0.0025, crs = 4326) {
  dw <- cbind(c(x - w, x + w, x + w, x - w, x - w),
              c(y + w, y + w, y - w, y - w, y + w))
  st_sf(geometry = st_sfc(st_polygon(list(dw))), crs = crs)
}
```


# Label Uncertainty
```{r}
sites_mapped_gt2 <- ctlg %>%
  group_by(Class, name) %>%
  count() %>%
  ungroup() %>%
  filter(n > 2)
sites_mapped_gt2

multi_site_list <- lapply(1:100, function(x) {
#multi_site_list <- lapply(1:nrow(sites_mapped_gt2), function(x) {
  
  multi_site <- sites_mapped_gt2 %>% slice(x)

# Filter the associations for the selected site
  multi_site_assns <- ctlg %>% filter(name == multi_site$name)

# Create an empty raster with the same extent as the target site
  target <- multi_site_assns %>%
    slice(1) %>%
    mutate(polybox(x, y)) %>%
    st_as_sf()

  # Rasterize at 224x224 pixels
  r <- rast(ext(target), nrows = 224, ncols = 224, crs = "EPSG:4326")
  values(r) <- 0

  # Process each association
  multi_site_assnsr <- lapply(1:nrow(multi_site_assns), function(x) {
    assn <- multi_site_assns[x, ]
    if (assn$nflds > 0) {
      assn_flds <- flds %>%
        filter(assignment_id == assn$assignment_id)
      r <- rasterize(assn_flds, r, background = 0)
    } else {
      r
    }
    return(r)
  }) %>% do.call(c, .)
})
```

# Calculations
```{r}
thresholds <- seq(0, 1, 0.2)
reclmat <- data.frame(
  lower = thresholds[-length(thresholds)],
  upper = thresholds[-1],
  index = 1:5,
  class = c("No field, high confidence", "No field, medium confidence", "Uncertain", "Field, medium confidence", "Field, high confidence")
)


pfield_list <- lapply(1:length(multi_site_list), function(x) {
  app_result <- app(multi_site_list[[x]], sum)
  nlyr_result <- nlyr(multi_site_list[[x]])
  app_result / nlyr_result
})

conf_class_list <- lapply(1:length(pfield_list), function(x) {
  classify(pfield_list[[x]], rcl = reclmat[, -4], include.lowest = TRUE)
})

# levels() from pdf doesn't work for lapply but categories() does
conf_class_list <- lapply(1:length(conf_class_list), function(x) {
  categories(conf_class_list[[x]], value = reclmat[, 3:4])
})

#plot(conf_class_list[[9]], mar = c(0, 1, 0, 12))
```

# Calculations + Formatting
```{r, warning = FALSE}
conf_table <- lapply(1:length(conf_class_list), function(x) {
  freq(conf_class_list[[x]])
})
#conf_table[[9]]


nofld <- lapply(1:length(pfield_list), function(x) {
  mask(1 - pfield_list[[x]], pfield_list[[x]] < 0.5, maskvalue = 0, updatevalue = 0)
})
#nofld[[2]]

fld <- lapply(1:length(pfield_list), function(x) {
  mask(pfield_list[[x]], pfield_list[[x]] >= 0.5, maskvalue = 0, updatevalue = 0)
})
# mcLapply attempt below, did not help!
#fld <- mclapply(1:length(pfield_list), function(x) {
#  mask(pfield_list[[x]], pfield_list[[x]] >= 0.5, maskvalue = 0, updatevalue = 0)
#}, mc.cores = 4)
#fld[[2]]


fld_nofld_conf <- lapply(1:length(fld), function(x) {
  c(fld[[x]], nofld[[x]], fld[[x]] + nofld[[x]])
})
# mcLapply attempt below, did not help!
#fld_nofld_conf <- mclapply(1:length(fld), function(x) {
#  c(fld[[x]], nofld[[x]], fld[[x]] + nofld[[x]])
#}, mc.cores = 4)
#fld_nofld_conf[[9]]

fld_nofld_conf <- lapply(1:length(fld_nofld_conf), function(x) {
  names(fld_nofld_conf[[x]]) <- c("field agreement", "no field agreement", "overall agreement")
  return(fld_nofld_conf[[x]])
})
#fld_nofld_conf[[9]]


#plot(fld_nofld_conf[[9]], nr = 1)

#global(fld_nofld_conf[[9]][[3]], mean)
```

# Example Site
```{r}
set.seed(9)

#insight to the counts for the site
conf_table[[9]]

plot(fld_nofld_conf[[9]], nr = 1)
```

For this site, we see that the field agreement averaged to:
```{r}
global(fld_nofld_conf[[9]][[1]], mean)
```

The no field agreement averaged to:
```{r}
global(fld_nofld_conf[[9]][[2]], mean)
```

It is important to understand what the percentages mean. 
0 represents no field. 1 represents a field.
The closer to 0, the more overall agreement on land not being a field.

The closer to 1, the more overall agreement on land being a field.

Knowing the average overall agreement does not provide as much insight as usual
averages would. Due to the data, knowing where the average overall agreement 
lies on a 0-1 scale simply informs if the site tends to have more crop fields
than non-crop fields. i.e.,
```{r}
global(fld_nofld_conf[[9]][[3]], mean)
```
Above 0.5 means the site is likely to be mostly crop fields, vice-versa if below.



# Summary Analysis
```{r}
all_site_fldmeans <- lapply(1:length(fld_nofld_conf), function(x) {
  global(fld_nofld_conf[[x]][[1]], mean)
})

all_site_nofldmeans <- lapply(1:length(fld_nofld_conf), function(x) {
  global(fld_nofld_conf[[x]][[2]], mean)
})

overall_dataset_mean <- mean(unlist(all_site_means))
overall_dataset_mean
# explain the range.

boxplot(unlist(all_site_means))
# do a boxplot for fields, for non-fields and their agreement for each respectively
# how hard it is to determine a field vs non-field

fld[[3]]
```

